---
title: "DA 3"
author: "Khawaja Hassan"
date: "1/23/2022"
output: pdf_document
---



```{r setup , echo=FALSE,include=FALSE}

rm(list=ls())
#####################
## Import libraries##
#####################
install.packages("tinytex")
library(tidyverse)
# library(lspline)
# library(cowplot)
# library(boot)
library(estimatr)
# library(huxtable)
# library(stargazer)
library(modelsummary)
# library(kableExtra)
# library(dplyr)
library(data.table)
library(fixest)
library(caret)
# library(skimr)
# library(grid)
# library(glmnet)
library(gridExtra)

```

```{r, echo=FALSE,include=FALSE}
###########################################
## Set data  functions##
###########################################

# Importing the data
data_org <- fread("https://osf.io/4ay9x/download")



```

```{r,echo=FALSE, warning=FALSE}

################
# Data Munging #
################

# Counting Number of Observation in each Occupation 
a <- data_org[,.(Count=.N),by=occ2012]

# Filtering the data for the chosen occupation (Secondary-Teacher)
dt <- data_org %>% filter(occ2012==2320)
rm(data_org)
dt <-  data.table(dt)

# Creating a new wage per hour variable w
dt <- dt[, w := earnwke/uhours]

# Filtering on uhours to work on full time employee (minimum 40 hours)/181 employee dropped since they were working less than 40 hours 

dt <- dt[uhours >= 40]

# counting Number observation in each Education Level
# Filter:Bachelor's degree or Higher (dropping 11 observation )

dt[,.(count=.N), by= grade92]
dt <- dt[grade92 >= 43]

# Checking age greater than 20/ Min age is 22 with sufficient observation
b <- dt[,.(count=.N), by= age]

# Looking at a quick summary of some of the key variables
datasummary(w + grade92 + age ~ Mean + SD + Min + Max + P25 + P75 + N , data = dt)

# Finding missing values

to_filter <- sapply(dt, function(x) sum(is.na(x)))
to_filter[to_filter > 0]

# dropping the ethnic column since there are 919 NA observation in the column 
dt <- dt[, ethnic := NULL]


```


```{r,echo=FALSE}
# Creating levels for grade92 variable in a new variable educ#
dt <- dt[grade92 == 43, educ := "Bachelors"]
dt <-dt[grade92 == 44, educ := "Masters"]
dt <-dt[grade92 == 45, educ := "Professional-School"]
dt <-dt[grade92 == 46, educ := "Doctorate degree"]

dt <- dt[, grade92 := NULL]

# Creating age squared column

dt <- dt[, agesq := age^2]

# checking result
datasummary(w*factor(educ) + age ~ Mean + SD + Min + Max + P25 + P75 + N , data = dt)

# Race

datasummary( w*factor(race) ~ N + SD +Percent() + Mean, data = dt )
# creating binary for white and non white
dt$white <- ifelse(dt$race==1,1,0)
dt <- dt %>%
  select(-c(race))

# Creating factor sex variable

dt <- dt[sex == 1, gender := "male"]
dt <- dt[sex == 2, gender := "female"]
dt <- dt %>%
  select(-c(sex))

# Creating factor variable for marital variable

dt[marital <= 2, married_status := "married"]
dt[marital <= 6 & marital >= 3, married_status := "separated"]
dt[marital == 7, married_status := "never married"]
dt <- dt %>%
  select(-c(marital))


# Ownchild

dt[,.(count=.N), by= ownchild]

# Create if individuals own child or no
  dt <- dt %>% mutate(ownchild=case_when(
  ownchild==0 ~ 0,
  TRUE ~ 1))
 datasummary( w*factor(ownchild) ~ N + SD + Mean, data = dt ) 
  
# dropping as it is the same as own child
dt <- dt %>%
  select(-chldpres)

# ind02( only elementary school)
datasummary( w*(ind02) ~ N + Percent() + Mean, data = dt ) 
# exclude from model
dt <- dt %>% select(-ind02)

# class
datasummary( w*factor(class) ~ N + Percent() + Mean, data = dt ) 
# include as factor
dt <- dt[class == "Government - Federal"| class=="Government - Local"|class=="Government - State", Sector := "Government"]
dt <- dt[class == "Private, For Profit"| class=="Private, Nonprofit", Sector := "Private"]

datasummary( w*factor(Sector) ~ N + Percent() + Mean, data = dt ) 
dt <- dt %>% select(-class)


# unionmme
datasummary( w*factor(unionmme) ~ N + Percent() + Mean, data = dt ) 

# state 
datasummary( w*factor(state) ~ N + Percent() + Mean, data = dt ) 

# lfsr94(mean difference is minor between these two so excluding from)
datasummary( w*factor(lfsr94) ~ N + Percent() + Mean, data = dt ) 
# exclude from model
dt <- dt %>% select(-lfsr94)

# prcitshp
datasummary( w*factor(prcitshp) ~ N + Percent() + Mean, data = dt ) 

dt <- dt[prcitshp=="Native, Born Abroad Of US Parent(s)"|prcitshp=="Native, Born in PR or US Outlying Area"|prcitshp=="Native, Born In US",origin := "Native"]
dt <- dt[prcitshp=="Foreign Born, Not a US Citizen"|prcitshp=="Foreign Born, US Cit By Naturalization",origin := "Foreign "]

dt <- dt %>% select(-prcitshp)

datasummary( w*factor(origin) ~ N + Percent() + Mean, data = dt ) 

##Creating region based on states, Used Aftabs ideas to group these codes in region 

dt <- dt[stfips %in% c("WA", "OR", "MT", "ID", "WY", "NV", "UT", "CO", "AZ", "NM", "HI", "AK", "CA"), region := "west"]
dt <- dt[stfips %in% c("ND", "SD", "NE", "KS", "MN", "IA", "MO", "WI", "IL", "IN", "MI", "OH"), region := "mid-west"]
dt <- dt[stfips %in% c("OK", "TX", "AR", "LA", "KY", "TN", "MS", "AL", "WV", "VA", "NC", "SC", "GA", "FL", "DC","MD","DE"), region := "south"]
dt <- dt[stfips %in% c("PA", "NY", "VT", "NH", "ME","MA","RI","CT","NJ"), region := "north-east"]

# excluding unwanted variable
dt <- dt %>% select(-c(V1,hhid,intmonth,weight,occ2012,stfips,state))

```

```{r, echo=FALSE}
###############################
##Setting models interactions##
###############################

#1#
datasummary( w*factor(white)*gender ~ N + Percent() + Mean, data = dt ) 
# wage is  very different based on white and gender, so  interaction needed

#Boxplot#
white_gender <- ggplot(dt, aes(x = factor(white), y = w,
                              fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Race",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))

#2#
datasummary( w*factor(educ)*gender ~ N + Percent() + Mean, data = dt )
# wage is very different based on education and gender

## Boxplot##
educ_gender <- ggplot(dt, aes(x = factor(educ), y = w,
                              fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Education",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))


#3#
datasummary( w*married_status*gender ~ N + Percent() + Mean, data = dt )#3
# wage is very different based on , married_status and gender

## Boxplot##
Status_gender <- ggplot(dt, aes(x = factor(married_status), y = w,
                              fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Married status",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))

#4#

datasummary( w*Sector*gender ~ N + Percent() + Mean, data = dt )#4
# wage is very different based on education and gender

## Boxplot##
Sector_gender <- ggplot(dt, aes(x = factor(Sector), y = w,
                              fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Sector",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))


#5#
datasummary( w*unionmme*gender ~ N + Percent() + Mean, data = dt ) #4


## Boxplot##
Union_gender <- ggplot(dt, aes(x = factor(unionmme), y = w,
                              fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Union Membership",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))

#6#
datasummary( w*unionmme*Sector ~ N + Percent() + Mean, data = dt )#3

## Boxplot##
Union_Sector <- ggplot(dt, aes(x = factor(unionmme), y = w,
                              fill = factor(Sector), color=factor(Sector))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Union Membership",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))

#7#
datasummary( w*origin*Sector ~ N + Percent() + Mean, data = dt ) # not including since the wage differential is low
 #wage is not very different

#8#
datasummary(w*Sector*region ~ N + Percent() + Mean, data = dt)#4

## Boxplot##
Sector_Region<- ggplot(dt, aes(x = factor(region), y = w,
                              fill = factor(Sector), color=factor(Sector))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Region",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))

#9#
datasummary(w*factor(white)*factor(married_status)  ~ N + Percent() + Mean, data = dt )#4

## Boxplot##
White_Status<- ggplot(dt, aes(x = factor(married_status), y = w,
                              fill = factor(white), color=factor(white))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Marriage Status",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))


#10#
datasummary(w*factor(educ)*region  ~ N + Percent() + Mean, data = dt )#4
#11#
datasummary(w*origin*factor(educ)  ~ N + Percent() + Mean, data = dt) 
# not including since the wage differential is minor and number of observation for Professional school is only 1

#12#
datasummary(w*factor(ownchild)*factor(gender) ~ N + Percent() + Mean, data = dt )
# not including since the wage differential is minor

#13#
datasummary(w*factor(white)*factor(educ)  ~ N + Percent() + Mean, data = dt )#3

white_educ<- ggplot(dt, aes(x = factor(educ), y = w,
                              fill = factor(white), color=factor(white))) +
  geom_boxplot(alpha=0.8, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_manual(name="",
                     values=c('red','blue')) +
  scale_fill_manual(name="",
                    values=c('red','blue')) +
  labs(x = "Education Level",y = "Wage per Hour (USD)")+
  scale_y_continuous(expand = c(0.01,0.01), limits=c(0, 70), breaks = seq(0,70, 10))+
  ggthemes::theme_economist() +
  theme(legend.position = c(0.15,0.85), axis.text.x = element_text(angle=45, vjust=.5))
```

```{r,echo=FALSE,include=FALSE}
###################################
## Setting up Regressions models ##
###################################

model1 <- as.formula(w ~ educ)

model2 <- as.formula(w ~ educ+age+agesq)

model3 <- as.formula(w ~ educ+age+agesq+region+ownchild+married_status+white+factor(white)*gender+married_status*gender+unionmme*Sector+unionmme*gender+white*educ)

model4 <- as.formula(w ~ educ+age+agesq+region+ownchild+married_status+white+ownchild+factor(white)*gender+married_status*gender+unionmme*Sector+unionmme*gender+white*educ+factor(educ)*region+factor(white)*factor(married_status)+Sector*region+unionmme*gender+Sector*gender)


### Running the regressions
reg1 <- feols(model1, data = dt , vcov="hetero")
reg2 <- feols(model2, data = dt , vcov="hetero")
reg3 <- feols(model3, data = dt , vcov="hetero")
reg4 <- feols(model4, data = dt , vcov="hetero")


 # evaluation of the models: using all the sample#
fitstat_register("k", function(x){length( x$coefficients ) - 1}, "No. Variables")           
etable( reg1 , reg2 , reg3 , reg4 , fitstat = c('aic','bic','rmse','r2','n','k'), keepFactors = TRUE )


#####################
# Cross-validation for better evaluation of predictive performance
# Simple k-fold cross validation setup:
# 1) Used method for estimating the model: "lm" - linear model (y_hat = b0+b1*x1+b2*x2 + ...)
# 2) set number of folds to use (must be less than the no. observations)
k <- 4

# We use the 'train' function which allows many type of model training -> use cross-validation
set.seed(200)
cv1 <- train(model1, dt, method = "lm", trControl = trainControl(method = "cv", number = k))

# Check the output:
cv1
summary(cv1)
cv1$results
cv1$resample

set.seed(200)
cv2 <- train(model2, dt, method = "lm", trControl = trainControl(method = "cv", number = k))
set.seed(200)
cv3 <- train(model3, dt, method = "lm", trControl = trainControl(method = "cv", number = k), na.action = "na.omit")
set.seed(200)
cv4 <- train(model4, dt, method = "lm", trControl = trainControl(method = "cv", number = k), na.action = "na.omit")

# Calculate RMSE for each fold and the average RMSE as well
cv <- c("cv1", "cv2", "cv3", "cv4")
rmse_cv <- c()


for(i in 1:length(cv)){
  rmse_cv[i] <- sqrt((get(cv[i])$resample[[1]][1]^2 +
                        get(cv[i])$resample[[1]][2]^2 +
                        get(cv[i])$resample[[1]][3]^2 +
                        get(cv[i])$resample[[1]][4]^2)/4)
}


# summarize results
cv_mat <- data.frame(rbind(cv1$resample[4], "Average"),
                     rbind(cv1$resample[1], rmse_cv[1]),
                     rbind(cv2$resample[1], rmse_cv[2]),
                     rbind(cv3$resample[1], rmse_cv[3]),
                     rbind(cv4$resample[1], rmse_cv[4])
)

colnames(cv_mat)<-c("Resample","Model1", "Model2", "Model3", "Model4")
cv_mat 

# Show model complexity and out-of-sample RMSE performance
m_comp <- c()
models <- c("reg1", "reg2", "reg3", "reg4")
for( i in 1 : length(cv) ){
  m_comp[ i ] <- length( get( models[i] )$coefficient  - 1 ) 
}

m_comp <- tibble( model = models , 
                  complexity = m_comp,
                  RMSE = rmse_cv )

kfold <- ggplot( m_comp , aes( x = complexity , y = RMSE ) ) +
  geom_point(color='red',size=2) +
  geom_line(color='blue',size=0.5)+
  scale_y_continuous(limits = c(10.8,11.7))+
  labs(x='Number of explanatory variables',y='Averaged RMSE on test samples',
       title='Prediction performance and model compexity') +
  ggthemes::theme_economist_white()

# plotting results
line <- ggplot(dt, aes(x=predict(reg2, dt), y=w)) + 
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, size = 0.5) +
  scale_x_continuous(limits = c(0,40)) + 
  scale_y_continuous(limits = c(0,60)) +
  ggthemes::theme_economist_white()

```

```{r}

## Arranging interaction plots into a grid for better presentation

grid.arrange(Sector_gender, Sector_Region, Status_gender, Union_gender,
             nrow = 2, ncol = 2)

grid.arrange( Union_Sector,white_gender,White_Status, nrow = 2, ncol = 2)


```

## Introduction ##
 
The purpose of this project was to run analysis and prediction of [Occupational& Earning data set](https://osf.io/g8p9j/) with Regression Analysis. The ultimate goal was to optimize prediction model for the hourly earning based on a particular occupation to capture their interaction with the predictors. The occupation selected for our analysis was Secondary school teachers and their codes are provided in on the following [link](https://osf.io/57n9q/). 

## Data Cleaning & Munging ##

The choice of occupation for our model was *Secondary Teacher* and the number observation after we filter it came around was 1100. However, before the data had to be modified before we could actually run our regression model and make our predictive analysis. The data cleaning was divided into three major parts filtering, adding categorical & binary variables and dropping columns which were not relevant or had missing values. For filtering, we included only those people whose minimum hour were greater than 40 to account for only full-time employees and dropped observation whose education level was less Bachelor Degree.Secondly, for the variables like married status, gender, class , state ,citizenship status we group them and made new categorical variable since the variation between the given categories were not that significant. Lastly, we dropped ethnic column since more than 90 percent of observation were missing , along with we removed the columns that were not used in our models.

## Variable Interaction ##

Before we could start running regression model we had to account for the interaction between the variable and decide which predictor to choose. The base for selection of predictor was if the wage differential was greater than $2 and the number of observation were sufficient enough to make a comparison. Since we are looking at hourly wage even 2 dollar difference would make a huge difference if we compare weekly or the monthly wage. However, with some interaction like one with the origin and education the number of observation in were below 5 and thus we dropped the those interaction in the model. Moreover, there are several interaction we did run but did not added it in our model because there result were insignificant.

## Regression Model

We build 4 predictive model starting from the basic and made our model more and more complex by adding predictors. Now once we run our model we had to validate their performance based on the following criteria

1- RMSE in the full sample
2- Cross-validated RMSE 
3- BIC in the full sample

The result from the regression table showed that Model 4 had the lowest RMSE of 10.55, although in comparison to Model 3 the 





